\documentclass[11pt]{article}

\usepackage{tgpagella}
\linespread{1.0}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{hyperref}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{tikz-cd}
\usepackage{color}

\theoremstyle{definition}
\newtheorem{prob}{Problem}

\newcommand{\kk}[1]{\mathbf{#1}}
\newcommand{\cc}[1]{\mathcal{#1}}

\def\<{\langle}
\def\>{\rangle}

\DeclareMathOperator{\supp}{supp}

\def\head{
\begin{center}
\textbf{\LARGE Homework set 2}
\\[3pt]
Due by 15:00 on Monday, September 4, 2023.
\end{center}
\medskip
}

\begin{document}

\head

Please select three problems to solve and hand in written solutions either in person or to \verb+gunnar@magnusson.io+.

A \emph{metrizable vector space} is a vector space $V$ that is
equipped with a metric $d$ for which the addition map $(x, y) \mapsto x +
y$ and multiplication map $(\lambda, x) \mapsto \lambda x$ are continuous.
We suppose here that $V$ is defined over a field $k$ such that $\kk R \subset k$
(for example, the real or complex numbers\footnote{Or the field $\kk C(X)$ of rational functions over $\kk C$ if you want to get weird.}).


\begin{prob}
Let $T_a(x) = x + a$ and $M_\lambda(x) = \lambda x$ for $a \in V$ and $\lambda
\in k$.
Show that $T_a$ and $M_\lambda$ are homeomorphisms of $V$ with itself; that is,
they are continuous maps that have a continuous inverse.
\end{prob}


\begin{prob}
Let $V$ and $W$ be metrizable vector spaces and $f : V \to W$ a linear map.
Show that $f$ is continuous if and only if it is continuous at $0$.
\end{prob}


A subset $E \subset V$ of a metrizable vector space is \emph{bounded} if for
any neighborhood $U$ of $0$ there exists a $\lambda > 0$ such that $\mu E
\subset U$ for any $\mu \geq \lambda$.

Note that this is \emph{not} the same notion of boundedness we get from the
metric on $V$; it can be defined if we only have a topology and not a metric.
Sets can be bounded in one notion and not the other.


\begin{prob}
Let $U$ be an open set that contains $0$.
Let $(r_j)$ be an increasing sequence of positive real numbers such that $r_j
\to \infty$. Show that $V = \bigcup_{j=1}^\infty r_j U$.
Conclude that if $K \subset V$ is compact, then $K$ is bounded.
\end{prob}

\begin{prob}
Let $E \subset V$ be a set.
Show that the following are equivalent:
\begin{enumerate}
\item
$E$ is bounded.

\item
If $(x_n)$ is a sequence in $E$ and $(\lambda_n)$ is a sequence of scalars such
that $\lambda_n \to 0$ as $n \to \infty$, then $\lambda_n x_n \to 0$ as $n \to
\infty$.
\end{enumerate}
\end{prob}

A map $f : V \to W$ between metrizable vector spaces is \emph{bounded} if
it maps bounded sets to bounded sets.

\begin{prob}
Let $V$ and $W$ be metrizable vector spaces and $f : V \to W$ a linear map.
Show that if $f$ is continuous then it is bounded.%
\footnote{The converse is true if we assume the metric is translation
invariant, that is, it satisfies $d(x + z, y + z) = d(x,y)$ for all $x,y,z \in
V$.}
Find spaces $V$ and $W$ and a linear function $f : V \to W$ that is not continuous.
\end{prob}

\end{document}
