\documentclass[11pt]{article}

\usepackage{tgpagella}
\linespread{1.0}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{hyperref}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{tikz-cd}
\usepackage{color}

\theoremstyle{definition}
\newtheorem{prob}{Problem}

\newcommand{\kk}[1]{\mathbf{#1}}
\newcommand{\cc}[1]{\mathcal{#1}}

\def\<{\langle}
\def\>{\rangle}

\DeclareMathOperator{\supp}{supp}

\def\head{
\begin{center}
\textbf{\LARGE Homework set 3}
\\[3pt]
Due by 15:00 on Monday, September 11, 2023.
\end{center}
\medskip
}

\begin{document}

\head

Please select \textbf{three} problems to solve and hand in written solutions
either in person or to \verb+gunnar@magnusson.io+. You may quote problems from
older homework sets and results we've read in the textbook if you feel like
they help.

Recall that in last week's problems we defined a subset $E \subset V$ of a
metrizable vector space to be \emph{bounded} if for any neighborhood $U$ of $0$
there exists a $\lambda > 0$ such that $E \subset \mu U$ for any $\mu \geq
\lambda$.


\begin{prob}
Let $V$ be a normed vector space.
Show that $E \subset V$ is bounded in the metrizable sense
if and only if it is bounded in the usual sense, that is, there exists
$r > 0$ such that $|x| \leq r$ for all $x \in E$.
\end{prob}

\begin{prob}
Let $V$ be finite-dimensional and let
$| \, \cdot \, |_1$ and
$| \, \cdot \, |_2$ be two norms on $V$.
Show that the norms are equivalent, meaning that there exist constants $c$ and
$C$ such that
\[
c |x|_1 \leq |x|_2 \leq C|x|_1
\]
for all $x \not= 0$.
\end{prob}

A set $E \subset V$ is \emph{convex} if for any $x, y \in E$ the vectors
$\lambda x + (1 - \lambda) y$ are also in $E$ for $0 \leq \lambda \leq 1$.

\begin{prob}
\begin{enumerate}
\item
Show that the unit ball in a normed vector space is convex.

\item
Show that a linear subspace $E \subset V$ is convex.

\item
If $f : V \to W$ is linear and $E \subset V$ is convex, show that $f(E)$ is convex.
\end{enumerate}
\end{prob}

\begin{prob}
Let $V$ and $W$ be normed vector spaces and let $f : V \to W$ be linear.
Show that $f$ is continuous if and only if it is bounded, in the sense
that there exists a constant $C > 0$ such that $|f(x)| \leq C |x|$ for all $x$.
\end{prob}

\begin{prob}
Let $V$ be a normed vector space.
We are going to prove that $V$ can be embedded in a Banach space.
\begin{enumerate}
\item
Define $X$ to be the set of Cauchy sequences in $V$.
Show that $X$ may be given the structure of a vector space.

\item
Show that $(x_n) \mapsto \lim_{n \to \infty} |x_n|$ defines a seminorm on $X$;
that is a function that satisfies the conditions to be norm except $|x| = 0$
does not imply $x = 0$.

\item
Show that $N := \{ x \in X \mid |x| = 0 \}$ is a closed subspace of $X$, and that
the seminorm on $X$ induces a norm on the quotient space $X / N$.

\item
Show that $X / N$ is complete, and thus a Banach space.

\item
Show that the map $f : V \to X / N$ that sends $x$ to the image of the sequence
$(x,x,\ldots)$ under the quotient map is linear, injective, and continuous.

\item
Show that $X / N$ satisfies the following universal property:
If $Y$ is a Banach space and $f : V \to Y$ is a continuous linear map, then
there is a unique continuous linear map $\hat f : X/N \to Y$ such that the
following diagram commutes:
\[
\begin{tikzcd}
V \ar[dr,"f"] \ar[d] &
\\
X/N \ar[r,"\hat{f}"] & Y
\end{tikzcd}
\]
\end{enumerate}
\end{prob}

\end{document}
