\documentclass[11pt]{article}

\usepackage{tgpagella}
\linespread{1.0}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{hyperref}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{tikz-cd}
\usepackage{color}

\theoremstyle{definition}
\newtheorem{prob}{Problem}

\newcommand{\kk}[1]{\mathbf{#1}}
\newcommand{\cc}[1]{\mathcal{#1}}

\def\<{\langle}
\def\>{\rangle}

\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\im}{im}

\def\head{
\begin{center}
\textbf{\LARGE Homework set 4}
\\[3pt]
Due by 15:00 on Monday, September 18, 2023.
\end{center}
\medskip
}

\begin{document}

\head

Please select \textbf{three} problems to solve and hand in written solutions
either in person or to \verb+gunnar@magnusson.io+. You may quote problems from
older homework sets and results we've read in the textbook if you feel like
they help.

If $f : V \to W$ is a linear map we denote its kernel by $\ker f = \{ x \in V
\mid f(x) = 0 \}$.

\begin{prob}
Let $f : V \to V$ be a linear map.
\begin{enumerate}
\item
Suppose $V$ is finite-dimensional.
Show that $f$ is bijective if and only if $\ker f = 0$.

\item
Consider $l^\infty$, the space of bounded sequences $(x_n)$ of real
numbers.
Let $f : l^\infty \to l^\infty$ be the right shift operator $f(x_n) =
(0,x_1,x_2,\ldots)$.
Show that $\ker f = 0$ but that $f$ is not bijective.
\end{enumerate}
\end{prob}

\begin{prob}
Let $f : V \to W$ be a surjective bounded operator from a normed space to a
normed space. Suppose that there is a $c > 0$ such that $|f(x)| \geq c|x|$ for
all $x$. Show that $f^{-1} : W \to V$ exists and is bounded.
\end{prob}

\begin{prob}
Let $l^\infty$ be the space of bounded sequences $(x_n)$ of real numbers.
\begin{enumerate}
\item
Show that $f : l^\infty \to l^\infty$ defined by $f(x_n) = (x_n / n)$ is linear
and bounded.

\item
Show that $f^{-1}$ exists and is linear but not bounded.
\end{enumerate}
\end{prob}

\begin{prob}
Let $V = C^0([0,1])$ be the space of real-valued continuous functions on $[0,1]$.
Let $T : V \to \kk R$ be defined by $T(f) = f(0)$.
\begin{enumerate}
\item
Show that $T$ is a linear functional.

\item
Consider the norm $|f| = (\int_0^1 |f(x)|^p dx)^{1/p}$ for $1 \leq p < \infty$.
Show that $T$ is not bounded with respect to this norm.

\item
Consider the norm $|f| = \sup_{0 \leq x \leq 1} |f(x)|$.
Show that $T$ is bounded with respect to this norm and calculate its norm.
\end{enumerate}
\end{prob}

\end{document}
