\documentclass[11pt]{article}

\usepackage{tgpagella}
\linespread{1.0}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{hyperref}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{tikz-cd}
\usepackage{color}

\newtheorem*{theo}{Theorem}
\theoremstyle{definition}
\newtheorem{prob}{Problem}
\newtheorem*{defi}{Definition}

\newcommand{\kk}[1]{\mathbf{#1}}
\newcommand{\cc}[1]{\mathcal{#1}}

\def\<{\langle}
\def\>{\rangle}

\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\im}{im}

\def\head{
\begin{center}
\textbf{\LARGE Homework set 5}
\\[3pt]
Due by 15:00 on Monday, September 25, 2023.
\end{center}
\medskip
}

\begin{document}

\head

Please select \textbf{three} problems to solve and hand in written solutions
either in person or to \verb+gunnar@magnusson.io+. You may quote problems from
older homework sets and results we've read in the textbook if you feel like
they help.

Recall that the (algebraic) dual of a vector space $V$ is the space of
linear maps $V^* = \Hom(V, \kk R)$.
For a normed space, we will usually want to talk about the bounded linear
functionals on the space.
We will denote the set of those by
$V^\vee := \{ f \in V^* \mid \text{$f$ is bounded} \}$.

\begin{prob}
If $V$ is a normed space and $S \subset V$ a subset, we define the distance from
a point $x$ to $S$ by $d(x, S) = \inf_{y \in S} |x - y|$.
Let $f : V \to \kk R$ be a bounded linear functional and let
\[
H_f = \{ x \in V \mid f(x) = 1 \}.
\]
Show that
\[
|f| = \frac{1}{d(0, H_f)}.
\]
\end{prob}

\begin{prob}
Let $V = \kk R^2$ and let $f : V \to V$ be the operator defined by the matrix
\[
A = \begin{pmatrix}
1 & 2 \\ 3 & 4
\end{pmatrix}.
\]
Calculate $|A|$ when $V$ has each of the norms below:
\begin{enumerate}
\item
$|(x,y)|_1 = |x| + |y|$.

\item
$|(x,y)|_2 = \sqrt{x^2 + y^2}$.

\item
$|(x,y)|_\infty = \max\{|x|, |y|\}$.
\end{enumerate}
If this is too difficult we may also accept bounds in terms of the other norms
or numerical approximations.
\end{prob}

\begin{prob}
Let $V$ and $W \not= 0$ be normed spaces.
Suppose that $\dim V = \infty$.
Show that there exists an unbounded linear operator $f : V \to W$.
Conclude that $V^\vee \not= V^*$.
(Hint: Every vector space has a basis.)
\end{prob}


If $X$ is a set and $f$ a function on a subset $S \subset X$, we say that
a function $\hat f$ on $X$ is an extension of $f$ if $\hat f(x) = f(x)$ for
all $x \in S$.


\begin{theo}[Hahn--Banach]
Let $V$ be a normed space and let $S \subset V$ be a subspace.
If $f \in S^\vee$ there exists a bounded extension $\hat f \in V^\vee$ of $f$
such that $|\hat f| = |f|$.
\end{theo}

\begin{prob}[Proof of Hahn--Banach, part 1]
Let $V$ be a normed space.
\begin{enumerate}
\item
Let $S \subset V$ be a linear subspace.
Let $x_1 \in V \setminus S$ and let $S_1 = S + \kk R x_1$.
Show that $S_1$ is a linear subspace that contains $S$, and that every vector
in $S_1$ can be written uniquely as $x = x_0 + \lambda x_1$, where $x_0 \in S$
and $\lambda \in \kk R$.

\item
Let $f \in S^\vee$.
Show that $f_1(x) = f(x_0)$ is a bounded extension of $f$ from $S^\vee$ to
$S_1^\vee$, and that $|f_1| = |f|$.
\end{enumerate}
\end{prob}

\emph{A partial order} on a set $S$ is a binary relation $\prec$
on $S$ that satisfies:
\begin{itemize}
\item
Reflexivity: $x \prec x$.

\item
Antisymmetry:
$x \prec y$ and $y \prec x$ imply $x = y$.

\item
Transitivity:
$x \prec y$ and $y \prec z$ imply $x \prec z$.
\end{itemize}
As an example, consider the inclusion $U \subset V$ of subsets of $S$.

A subset $T \subset S$ is totally ordered if for every $x,y$ in $T$ we have
either $x \prec y$ or $y \prec x$.
An element $y$ is an upper bound for $T$ if $x \prec y$ for every $x \in T$.
Finally an element $y \in S$ is maximal if $x \prec y$ implies $x = y$.

Zorn's lemma says that if $(S,\prec)$ is a partially ordered set such that
every totally ordered subset contains an upper bound, then $(S,\prec)$ contains
at least one maximal element.


\begin{prob}[Proof of Hahn--Banach, part 2]
Let $V$ be a normed space and $S \subset V$ a subspace.
Let also $f \in S^\vee$ be a bounded linear functional.
Denote by $\cc L$ the set of all bounded extensions $(M, g)$ of $f$ to a
subspace $M$ such that $|g| = |f|$.

\begin{enumerate}
\item
Show that $(S, f) \in \cc L$, so it is not empty.

\item
We write $(M,g) \prec (M',g')$ if $M \subset M'$ and $g'(x) = g(x)$ for all $x
\in M$.
Show that this is a partial order on $\cc L$.

\item
Suppose that $\cc F$ is a totally ordered subset of $\cc L$.
Define a set $W = \bigcup_{(M,f) \in \cc F} M$.
Show that $W$ is in fact a vector subspace of $V$.

\item
Suppose that $\cc F$ is a totally ordered subset of $\cc L$.
Define $W$ as above and define $h : W \to \kk R$ by
\(
h(x) = g(x)
\)
for any $(M,g)$ such that $x \in M$.
Show that this is well-defined, and that $h$ is an extension of $f$.

\item
Conclude that there exists a maximal extension $h : W \to \kk R$ of $f$,
and use part 1 to conclude that we must have $W = V$.
\end{enumerate}
\end{prob}

\end{document}
